### 2021-08-18 Paper_Higher_Education.Rmd
Add files via upload

---
title: "Influences on the access to higher education: Visualization and modeling for Chilean data"
author: 'Bastian Olea-Duran, Francisco A. Segovia, Nixon Jerez-Lillo, Eduardo Alarcón-Bustamante, Danilo Alvares'
date: "10/08/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#Setup
options(scipen = 999)
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(data.table)
library(bit64)
library(readstata13)
library(haven)
library(dplyr)
library(readr)
library(grid)
library(gridExtra)
library(factoextra)
library(chilemapas)
library(kableExtra)
library(rsconnect)
library(knitr)
library(lemon)
library(ggspatial)
library(sf)
library(ggsflabel)
library(grid)
library(gridExtra)
library(ggalluvial)
library(ggrepel)
options(digits = 3)
```

```{r Fig: 1, echo=FALSE, fig.height=7, fig.width=15, message=FALSE, warning=FALSE}
setwd("~/Documents/GitHub/HE_Chile")
db_final <- read_rds("BASE_PAPER.RDS")
db_final <- db_final %>% filter(JOVEN=="1")
db_final <- db_final %>% rename(COD_REG_ALU=codigo_region,Region=nombre_region)

matriculado_region <- db_final %>% 
  group_by(COD_REG_ALU, Region, MATRICULADO, TIPO_INST_1) %>% summarise(N=n()) #15297 students without region: only 0.64%

matriculado_region = dcast(matriculado_region,COD_REG_ALU + Region ~ TIPO_INST_1 + MATRICULADO)
matriculado_region <- matriculado_region %>% rename(Universidades=Universidades_1 , CFT=`Centros de Formación Técnica_1` ,IP=`Institutos Profesionales_1`)
matriculado_region <- matriculado_region %>% rename(No_Matriculados=NA_0)
matriculado_region <- matriculado_region %>% mutate(Total_Egresados=Universidades+CFT+IP+No_Matriculados)
matriculado_region <- matriculado_region %>% mutate(Total_Matriculados=Universidades+CFT+IP)
matriculado_region <- matriculado_region %>% filter(!is.na(COD_REG_ALU)) %>% filter(COD_REG_ALU>0)
matriculado_region <- matriculado_region %>% mutate(Tasa_Ingreso = Total_Matriculados/Total_Egresados*100)
matriculado_region <- matriculado_region %>% rename(codigo_region=COD_REG_ALU)

x = generar_regiones(mapa = chilemapas::mapa_comunas)
x$codigo_region = as.numeric(x$codigo_region)
x <- left_join(x, matriculado_region)
x <- x %>% mutate(Zoom=ifelse(codigo_region<8 | codigo_region==15 | codigo_region==16 | codigo_region==13,1,0))
x <- x %>% mutate(Porc = paste0(round(Total_Egresados/sum(Total_Egresados)*100,1),"%"))
x <- x %>% mutate(Region = ifelse(Region=="METROPOLITANA DE SANTIAGO", "RM",
                                  ifelse(Region=="VALPARAISO","VALPO",
                                         ifelse(Region =="BIOBIO", "BBIO",
                                                ifelse(Region=="LA ARAUCANIA", "ARAUC",
                                                       ifelse(Region=="LIBERTADOR GENERAL BERNARDO O'HIGGINS", "LGBO",
                                                              ifelse(Region=="LOS LAGOS","LAGOS",
                                                                     ifelse(Region=="COQUIMBO", "COQ",
                                                                            ifelse(Region=="ANTOFAGASTA","ANTOF",
                                                                                   ifelse(Region=="LOS RIOS", "RIOS",
                                                                                          ifelse(Region=="TARAPACA", "TPCA",
                                                                                                 ifelse(Region=="ATACAMA","ATCMA",
                                                                                                        ifelse(Region =="ARICA Y PARINACOTA", "AYP",
                                                                                                               ifelse(Region=="MAGALLANES Y DE LA ANTARTICA CHILENA","MAG", ifelse(Region=="AYSEN DEL GENERAL CARLOS IBANEZ DEL CAMPO", "AYSEN", Region)))))))))))))))

bar = ggplot(data=x, aes(x=reorder(factor(Region),Total_Egresados),y=Total_Egresados, fill=Total_Egresados))  + 
  coord_flip() + theme_minimal() + geom_bar(stat="identity") + labs(x="Region",y="Graduates") + 
  theme(legend.position = "none", axis.title = element_text(size=17), axis.text.y = element_text(size=14), 
        axis.text.x = element_text(size=14)) + 
  geom_text(aes(label = Porc), position = position_dodge(width= 1), vjust= 0, hjust = -0.05, size = 6, color="black") +
  scale_y_continuous(expand = expansion(mult = c(0, .3))) + 
  scale_fill_distiller(palette = "RdBu", direction=1, limits=c(13000,920000), breaks = seq(13000,920000,150000))

chile_1 = ggplot() +  geom_sf(data = x %>% filter(Zoom==1), aes(fill=Total_Egresados, geometry=geometry)) +  geom_sf_label_repel(data=x %>% filter(Zoom==1),aes(label = Region), nudge_x = -3, nudge_y = 0, seed = 10 ,colour="black", size=5) + coord_sf(xlim = c( -77.1,-65), ylim = c(-38, -17))
chile_1 = chile_1 + theme(legend.position="right", legend.box = "vertical") 
chile_1 = chile_1 + xlab("Longitude") + ylab("Latitude") + annotation_scale() + annotation_north_arrow(location='tr')
chile_1 = chile_1 + theme(plot.title = element_text(hjust = 0.5))  # + theme(legend.title=element_blank()) 
graf_total = chile_1 + theme_bw() + theme(legend.position = "none", axis.title = element_text(size=15)) +
  scale_fill_distiller(palette = "RdBu", direction=1, limits=c(13000,920000), breaks = c(13000,300000,600000,900000))

chile_2 = ggplot() +  geom_sf(data = x %>% filter(Zoom==0), aes(fill=Total_Egresados, geometry=geometry)) +  geom_sf_label_repel(data=x %>% filter(Zoom==0),aes(label = Region), nudge_x = 3, nudge_y = 0, seed = 10 ,colour="black", size=5) + coord_sf(xlim = c( -77.1,-65), ylim = c(-57, -38))
chile_2 = chile_2 + theme(legend.position="right", legend.box = "vertical") + labs(fill="Graduates")
chile_2 = chile_2 + xlab("Longitude") + ylab("Latitude") + annotation_scale() + annotation_north_arrow(location='tr')
chile_2 = chile_2 + theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(size=15), legend.text = element_text(size=15)) +scale_fill_distiller(palette = "RdBu", direction=1,limits=c(0,920000), breaks = seq(0,920000,150000))  # + theme(legend.title=element_blank()) 
graf_total2 = chile_2 + theme_bw() + theme(legend.position = "right", axis.title = element_text(size=15), legend.text = element_text(size=12), legend.title = element_text(size=15)) +scale_fill_distiller(palette = "RdBu",direction=1,limits=c(13000,920000), breaks = c(13000,300000,600000,900000))

Fig_1 <- grid.arrange(bar,graf_total,graf_total2, ncol=3, nrow=1)

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_1.eps",Fig_1, width=19.20, height=10.57, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_1.pdf",Fig_1, width=19.20, height=10.57, limitsize = FALSE)

```

```{r Fig: 2, 3 and 5, echo=FALSE, fig.align="center", fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
setwd("~/Documents/GitHub/HE_Chile")
db_final <- read_rds("BASE_PAPER.RDS")
db_final <- db_final %>% filter(JOVEN=="1")
db_final <- db_final %>% mutate(COD_ENSE=ifelse(COD_ENSE<400,"H-C","T-P")) #%>% filter(COD_ENSE=="H-C")
names(db_final) <- toupper(names(db_final))

colegios <- db_final %>% 
  group_by(RBD,COD_REG_RBD,COD_DEPE2,COD_COM_RBD,COD_ENSE,AGNO_EGRESO,MATRICULADO) %>% 
  summarise(N=n())

colegios = dcast(colegios,RBD + COD_REG_RBD + COD_DEPE2 + COD_COM_RBD + COD_ENSE + AGNO_EGRESO ~ MATRICULADO)
names(colegios)[7] = "No_Matriculado";names(colegios)[8] = "Matriculado"

colegios <- colegios %>% mutate(across(everything(), .fns = ~replace_na(.,0)))
colegios <- colegios %>% ungroup() %>% mutate(Tasa = (Matriculado)/(No_Matriculado+Matriculado)*100)
colegios <- colegios %>% ungroup() %>% mutate(Total_Egresados = Matriculado+No_Matriculado)

municipales <- colegios %>% filter(COD_DEPE2=="Municipal") %>% mutate(COD_DEPE2="Public")
municipales <- municipales %>% group_by(AGNO_EGRESO,COD_DEPE2) %>% summarise(No_Matriculado=sum(No_Matriculado),Matriculado=sum(Matriculado)) %>% ungroup() %>% mutate(Tasa=Matriculado/(Matriculado+No_Matriculado)*100)

subv <- colegios %>% filter(COD_DEPE2=="P.Sub") %>% mutate(COD_DEPE2="Semi-Private")
subv <- subv %>% group_by(AGNO_EGRESO,COD_DEPE2) %>% summarise(No_Matriculado=sum(No_Matriculado),Matriculado=sum(Matriculado)) %>% ungroup() %>% mutate(Tasa=Matriculado/(Matriculado+No_Matriculado)*100)

pagados <- colegios %>% filter(COD_DEPE2=="P.Pag") %>% mutate(COD_DEPE2="Private")
pagados <- pagados %>% group_by(AGNO_EGRESO,COD_DEPE2) %>% summarise(No_Matriculado=sum(No_Matriculado),Matriculado=sum(Matriculado)) %>% ungroup() %>% mutate(Tasa=Matriculado/(Matriculado+No_Matriculado)*100)

final <- bind_rows(municipales,subv,pagados)
final <- final %>% mutate(AGNO_EGRESO=AGNO_EGRESO+1) #Admission Process is 1 + Graduation Year
final <- final %>% ungroup() %>% mutate(Porc = paste0(round(Tasa,1),"%")) %>%
  mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")))

# Fig_5 -------------------------------------------------------------------

Fig_5 = ggplot(final, aes(x=AGNO_EGRESO, y=Tasa, group=factor(COD_DEPE2))) +
  geom_line(linetype = "3313")+
  geom_point(aes(shape=factor(COD_DEPE2)), size=4)+ 
  labs(x="Admission Process", y="Enrolment Rate", shape="Dependence") +
  coord_cartesian(xlim = c(2008,2020.2), ylim = c(0,90))+ scale_x_continuous(breaks = c(2007:2020)) +
  scale_y_continuous(breaks = seq(0,90,10), sec.axis = dup_axis()) +
  theme_bw() +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0), axis.title.x = element_text(size = 18, vjust = -0.5), 
        axis.text.y = element_text(size=13, angle = 0), axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=90), axis.title.y.right = element_blank(), 
        legend.position = "top") + 
  geom_vline(xintercept = 2015.8, linetype="dashed") + 
  annotate(geom = "text", x = 2015.6, y = 20, label = "Gratuidad", color = "black", angle = 90) + 
  geom_vline(xintercept = 2011.8, linetype="dashed") +
  annotate(geom = "text", x = 2011.6, y = 20, label = "St.Movement", color = "black", angle = 90) + 
  geom_vline(xintercept = 2009.8, linetype="dashed") + 
  annotate(geom = "text", x = 2009.6, y = 20, label = "LGE", color = "black", angle = 90) + 
  geom_vline(xintercept = 2006.8) + 
  geom_vline(xintercept = 2019.8, linetype="dashed") + 
  annotate(geom = "text", x = 2019.6, y = 20, label = "Soc.Outbreak", color = "black", angle = 90) + 
  geom_text(aes(label = Porc), vjust= -0.9, hjust = 0, size = 4, color="black", check_overlap = TRUE, nudge_y = 0.5) +
    scale_shape_manual(values = c('Public'=15, 'Semi-Private'=17, 'Private'=16))

Fig_5

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_5.eps",Fig_5, width=10.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_5.pdf",Fig_5, width=10.16, height=7.45, limitsize = FALSE)

# -----------------------------------------------------------------------------

final <- final %>% mutate(Total_Egresados = Matriculado+ No_Matriculado)
final <- final %>% filter(COD_DEPE2 !="All")
final <- final %>% ungroup() %>%  mutate(laggy = dplyr::lag(Total_Egresados, n = 1, default = NA))  
final <- final %>% mutate(Porc = ifelse(AGNO_EGRESO>2008, Total_Egresados/laggy*100-100,NA))
final <- final %>% ungroup() %>% mutate(Porc = ifelse(AGNO_EGRESO>2008,paste0(round(Porc,1),"%"),NA)) %>%
  mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")))

# Fig_2 -------------------------------------------------------------------

Fig_2 = ggplot(final, aes(x=AGNO_EGRESO, y=Total_Egresados, group=factor(COD_DEPE2))) +
  geom_line(linetype = "3313")+
  geom_point(aes(shape=factor(COD_DEPE2)), size=4)+ 
  labs(x="Admission Process", y="Graduates", shape="Dependence") +
  coord_cartesian(xlim = c(2008,2020.2), ylim = c(0,100000)) +
  scale_x_continuous(breaks = c(2007:2020)) + 
  scale_y_continuous(breaks = seq(0,100000,20000), sec.axis = dup_axis()) + 
  theme_bw() +
  theme(axis.title.y.right = element_blank(),plot.title = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=16, angle=0),legend.text=element_text(size=14, angle = 0),
        axis.title.x = element_text(size = 18, vjust=-0.5),   axis.text.y = element_text(size=15, angle = 0),
        axis.title.y = element_text(size=18, angle=90), axis.text.x = element_text(size=14, angle=90),
        legend.position = "top") +
  geom_vline(xintercept = 2015.8, linetype="dashed") + 
  annotate(geom = "text", x = 2015.6, y = 40000, label = "Gratuidad", color = "black", angle = 90) + 
  geom_vline(xintercept = 2011.8, linetype="dashed") + 
  annotate(geom = "text", x = 2011.6, y = 40000, label = "St.Movement", color = "black", angle = 90) + 
  geom_vline(xintercept = 2009.8, linetype="dashed") + 
  annotate(geom = "text", x = 2009.6, y = 40000, label = "LGE", color = "black", angle = 90) + 
  geom_vline(xintercept = 2006.8, linetype="dashed") + 
  geom_vline(xintercept = 2019.8, linetype="dashed") + 
  annotate(geom = "text", x = 2019.6, y = 40000, label = "Soc.Outbreak", color = "black", angle = 90) + 
  geom_text(aes(label = Porc), vjust= -0.9, hjust = 0, size = 4, color="black", check_overlap = TRUE, nudge_y = 0.5) +
  scale_shape_manual(values = c('Public'=15, 'Semi-Private'=17, 'Private'=16))

Fig_2

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_2.eps",Fig_2, width=10.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_2.pdf",Fig_2, width=10.16, height=7.45, limitsize = FALSE)

# -----------------------------------------------------------------------------

final <- final %>% group_by(AGNO_EGRESO) %>% mutate(Porcentaje = Total_Egresados/sum(Total_Egresados)*100)
final <- final %>% ungroup() %>% mutate(Porc = paste0(round(Porcentaje,1),"%"))
final <- final %>% group_by(AGNO_EGRESO) %>% mutate(ancho = sum(Total_Egresados))
final <- final %>% ungroup() %>% mutate(ancho = ancho / sum(ancho))

# Fig_3 -------------------------------------------------------------------

Fig_3 <- ggplot(final,aes(x = AGNO_EGRESO, y = Porcentaje, width = ancho, fill = COD_DEPE2)) +
  geom_col(colour = "black") +
  geom_text(aes(label = Porc), position = position_stack(vjust = 0.5)) + # if labels are desired
  facet_grid(~AGNO_EGRESO, scales = "free_x", space = "free_x") +
  scale_fill_brewer(palette = "Pastel1") +  
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        legend.position = "left", axis.title = element_blank()) +
   theme(legend.title=element_text(size=16, angle=0),legend.text=element_text(size=14, angle = 0),
        legend.position = "top", strip.text.x = element_text(size=14)) +
  labs(fill="Dependence") + scale_fill_grey(start=0.95, end=0.6)

Fig_3

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_3.eps",Fig_3, width=12.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_3.pdf",Fig_3, width=12.16, height=7.45, limitsize = FALSE)

# -----------------------------------------------------------------------------

```

```{r Fig: 4, echo=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=100), out.width = '55%'}
setwd("~/Documents/GitHub/HE_Chile")
db_final <- read_rds("BASE_PAPER.RDS")
db_final <- db_final %>% filter(JOVEN=="1")
db_final <- db_final %>% mutate(COD_ENSE=ifelse(COD_ENSE<400,"H-C","T-P")) #%>% filter(COD_ENSE=="H-C")
names(db_final) <- toupper(names(db_final))
db_final <- db_final %>% mutate(AGNO_EGRESO = AGNO_EGRESO+1)
db_final <- db_final %>% filter(GEN_ALU=="1" | GEN_ALU=="2") #1 observation drop
estudiantes <- db_final %>% mutate(GEN_ALU=ifelse(GEN_ALU==1, "Male","Female"))

GeomSplitViolin <- ggproto("GeomSplitViolin", GeomViolin, 
                           draw_group = function(self, data, ..., draw_quantiles = NULL) {
  data <- transform(data, xminv = x - violinwidth * (x - xmin), xmaxv = x + violinwidth * (xmax - x))
  grp <- data[1, "group"]
  newdata <- plyr::arrange(transform(data, x = if (grp %% 2 == 1) xminv else xmaxv), if (grp %% 2 == 1) y else -y)
  newdata <- rbind(newdata[1, ], newdata, newdata[nrow(newdata), ], newdata[1, ])
  newdata[c(1, nrow(newdata) - 1, nrow(newdata)), "x"] <- round(newdata[1, "x"])

  if (length(draw_quantiles) > 0 & !scales::zero_range(range(data$y))) {
    stopifnot(all(draw_quantiles >= 0), all(draw_quantiles <=
      1))
    quantiles <- ggplot2:::create_quantile_segment_frame(data, draw_quantiles)
    aesthetics <- data[rep(1, nrow(quantiles)), setdiff(names(data), c("x", "y")), drop = FALSE]
    aesthetics$alpha <- rep(1, nrow(quantiles))
    both <- cbind(quantiles, aesthetics)
    quantile_grob <- GeomPath$draw_panel(both, ...)
    ggplot2:::ggname("geom_split_violin", grid::grobTree(GeomPolygon$draw_panel(newdata, ...), quantile_grob))
  }
  else {
    ggplot2:::ggname("geom_split_violin", GeomPolygon$draw_panel(newdata, ...))
  }
})

geom_split_violin <- function(mapping = NULL, data = NULL, stat = "ydensity", position = "identity", ..., 
                              draw_quantiles = NULL, trim = TRUE, scale = "area", na.rm = FALSE, 
                              show.legend = NA, inherit.aes = TRUE) {
  layer(data = data, mapping = mapping, stat = stat, geom = GeomSplitViolin, 
        position = position, show.legend = show.legend, inherit.aes = inherit.aes, 
        params = list(trim = trim, scale = scale, draw_quantiles = draw_quantiles, na.rm = na.rm, ...))
}

estudiantes <- estudiantes %>% mutate(COD_DEPE2 = ifelse(COD_DEPE2=="Municipal","Public",
                                                     ifelse(COD_DEPE2=="P.Pag", "Private",
                                                            ifelse(COD_DEPE2=="P.Sub","Semi-Private",COD_DEPE2))))

estudiantes <- estudiantes %>% mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")))

Fig_4 = ggplot(data=estudiantes, aes(x=COD_DEPE2, y=NEM, fill=factor(GEN_ALU))) +
  geom_split_violin() + 
  theme_bw() + scale_fill_grey(start = 0.5, end = 0.7) +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0), axis.title.x = element_text(size = 18, vjust = -0.5), 
        axis.text.y = element_text(size=13, angle = 0), axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=15, angle=0), axis.title.y.right = element_blank(), 
        legend.position = "top") + 
  labs(x="Dependence",y="HGPA", fill="Sex") + 
  geom_boxplot(width=0.2,color="white",alpha=0.2) + 
  theme(legend.position = "top")

Fig_4

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_4.eps",Fig_4, width=12.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_4.pdf",Fig_4, width=12.16, height=7.45, limitsize = FALSE)

```

```{r Fig: 11, echo=FALSE, fig.align="center", fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
setwd("~/Documents/GitHub/HE_Chile")
db_final <- read_rds("BASE_PAPER.RDS")
db_final <- db_final %>% filter(JOVEN=="1")
db_final <- db_final %>% mutate(COD_ENSE=ifelse(COD_ENSE<400,"H-C","T-P")) #%>% filter(COD_ENSE=="H-C")
names(db_final) <- toupper(names(db_final))
db_final <- db_final %>% mutate(AGNO_EGRESO = AGNO_EGRESO+1,
                                PERCENTIL = ifelse(PERCENTIL<=50, "ClassRank 1-50%","ClassRank 51-100%" ))

colegios <- db_final %>% group_by(AGNO_EGRESO, COD_DEPE2,PERCENTIL,MATRICULADO) %>% summarise(N=n())

table(colegios$PERCENTIL)
colegios = dcast(colegios,AGNO_EGRESO + COD_DEPE2 + PERCENTIL ~ MATRICULADO)
names(colegios)[4] = "No_Matriculado";names(colegios)[5] = "Matriculado"

colegios <- colegios %>% mutate(across(everything(), .fns = ~replace_na(.,0)))
colegios <- colegios %>% ungroup() %>% mutate(Tasa = (Matriculado)/(No_Matriculado+Matriculado)*100)

colegios <- colegios %>% mutate(COD_DEPE2 = ifelse(COD_DEPE2=="Municipal","Public",
                                                     ifelse(COD_DEPE2=="P.Pag", "Private",
                                                            ifelse(COD_DEPE2=="P.Sub","Semi-Private",COD_DEPE2))))

colegios <- colegios %>% ungroup() %>% mutate(Porc = paste0(round(Tasa,1),"%"))
colegios <- colegios %>%  mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")))

Fig_11 = ggplot(colegios, aes(x=AGNO_EGRESO, y=Tasa, group=factor(COD_DEPE2))) +
  geom_line(linetype = "3313")+
  geom_point(aes(shape=factor(COD_DEPE2)), size=3)+ 
  labs(x="Admission Process", y="Enrolment Rate", shape="Dependence") +
  coord_cartesian(xlim = c(2008,2020.6), ylim = c(15,92))+
  scale_x_continuous(breaks = seq(2007,2020,1)) + 
  scale_y_continuous(breaks = seq(0,90,10), sec.axis = dup_axis()) + 
  theme_bw() +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),axis.title.x = element_text(size = 18, vjust = -0.5), 
        axis.text.y = element_text(size=12, angle = 0),axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=90), axis.title.y.right = element_blank(), strip.text = element_text(size = 16),
        legend.position="top") +
  #geom_text(aes(label = Porc), vjust= -0.9, hjust = 0, size = 2, color="black", check_overlap = TRUE, nudge_y = 0.5) + 
  facet_rep_wrap(~PERCENTIL, nrow=1,repeat.tick.labels = T) +
  scale_shape_manual(values = c('Public'=15, 'Semi-Private'=17, 'Private'=16))

Fig_11

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_11.eps",Fig_11, width=10.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_11.pdf",Fig_11, width=10.16, height=7.45, limitsize = FALSE)

```

```{r Fig: 6, echo=FALSE, fig.align="center", fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
setwd("~/Documents/GitHub/HE_Chile")
db_final <- read_rds("BASE_PAPER.RDS")
db_final <- db_final %>% filter(JOVEN=="1")

names(db_final) <- toupper(names(db_final))
db_final <- db_final %>% mutate(COD_ENSE=ifelse(COD_ENSE<400,"H-C","T-P")) #%>% filter(COD_ENSE=="H-C")
names(db_final) <- toupper(names(db_final))

db_final <- db_final %>% mutate(AGNO_EGRESO = AGNO_EGRESO+1)

db_final <- db_final %>% filter(GEN_ALU=="1" | GEN_ALU=="2")

colegios <- db_final %>% group_by(AGNO_EGRESO, GEN_ALU,MATRICULADO) %>% summarise(N=n())

colegios=dcast(colegios,AGNO_EGRESO + GEN_ALU ~ MATRICULADO)
names(colegios)[3] = "No_Matriculado";names(colegios)[4] = "Matriculado"

colegios <- colegios %>% mutate(across(everything(), .fns = ~replace_na(.,0)))
colegios <- colegios %>% ungroup() %>% mutate(Tasa = (Matriculado)/(No_Matriculado+Matriculado)*100)
colegios <- colegios %>% ungroup() %>% mutate(Porc = paste0(round(Tasa,1),"%"))

colegios <- colegios %>% mutate(GEN_ALU=ifelse(GEN_ALU==1, "Male","Female"))

Fig_6 = ggplot(colegios, aes(x=AGNO_EGRESO, y=Tasa, group=factor(GEN_ALU))) +
  geom_line(linetype = "3313") +
  geom_point(aes(shape=factor(GEN_ALU)), size=4) +
  theme() + 
  labs(x="Admission Process", y="Enrolment Rate", shape="Sex") +
  coord_cartesian(xlim = c(2008,2020), ylim = c(40,60))+ scale_x_continuous(breaks = seq(2007,2020,1)) +
  scale_y_continuous(breaks = seq(40,60,5), sec.axis = dup_axis()) + 
  theme_bw() + scale_color_grey(start = 0, end = .7) +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),axis.title.x = element_text(size = 18, vjust = -0.5),
        axis.text.y = element_text(size=12, angle = 0),axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=90), axis.title.y.right = element_blank(),
        legend.position="top") + 
   geom_text_repel(aes(label = Porc))+
   scale_shape_manual(values = c('Female'=16, 'Male'=17))

Fig_6

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_6.eps",Fig_6, width=12.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_6.pdf",Fig_6, width=12.16, height=7.45, limitsize = FALSE)

```

```{r Fig: 7, echo=FALSE, fig.align="center", fig.height=8, fig.width=12, message=FALSE, warning=FALSE}
setwd("~/Documents/GitHub/HE_Chile")

Alluvial <- read_rds("BASE_PAPER.RDS")
Alluvial <- Alluvial %>% filter(JOVEN=="1")
Alluvial <- Alluvial %>% rename(COD_REG_ALU=codigo_region,Region=nombre_region)

Alluvial <- Alluvial %>% mutate(COD_ENSE=ifelse(COD_ENSE<400,"Humanista-Científico","Técnico-Profesional"))
Alluvial <- Alluvial %>% filter(!is.na(COD_DEPE2),!is.na(COD_ENSE))

Alluvial <- Alluvial %>% mutate(TIPO_INST_3=ifelse(is.na(TIPO_INST_3),"Not-Enrolled",TIPO_INST_3))
Alluvial <- Alluvial %>% mutate(OECD_AREA=ifelse(is.na(TIPO_INST_3),"Not-Enrolled",OECD_AREA))
#Alluvial <- Alluvial %>% filter(MATRICULADO==1)
Alluvial <- Alluvial %>% group_by(COD_ENSE,COD_DEPE2,TIPO_INST_3) %>% summarise(N=n())

Alluvial <- Alluvial %>% mutate(COD_DEPE2 = ifelse(COD_DEPE2=="Municipal","Public",
                                                     ifelse(COD_DEPE2=="P.Pag", "Private",
                                                            ifelse(COD_DEPE2=="P.Sub","Semi-Private",COD_DEPE2))))

Alluvial <- Alluvial %>% ungroup() %>% mutate(TIPO_INST_3=ifelse(TIPO_INST_3=="Centros de Formación Técnica","PI or TTC",ifelse(TIPO_INST_3=="Centros de Formación Técnica Estatales","PI or TTC",ifelse(TIPO_INST_3=="Institutos Profesionales","PI or TTC",ifelse(TIPO_INST_3=="Universidades Estatales CRUCH", "CRUCH",ifelse(TIPO_INST_3=="Universidades Privadas","Non-CRUCH",ifelse(TIPO_INST_3=="Universidades Privadas CRUCH","CRUCH",TIPO_INST_3)))))))

Alluvial <- Alluvial %>%  mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")),
                                 TIPO_INST_3 = factor(TIPO_INST_3, levels = c("Not-Enrolled","CRUCH","Non-CRUCH","PI or TTC")))

Fig_7 <- ggplot(as.data.frame(Alluvial),
       aes(y = N, axis1 = COD_DEPE2, axis2=TIPO_INST_3)) +
  geom_alluvium(aes(fill = COD_DEPE2), width = 1/12) +
  geom_stratum(width = 1/12, fill = "black", color = "grey") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Dependence","Institution"), expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual", palette = "Set1") + labs(y="Graduates Enrolled", fill="Dependence") +
  theme_bw()+ theme(legend.position = "top",axis.title.x = element_text(size = 20, vjust = -0.5),
        axis.text.y = element_text(size=18, angle = 0),axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=20, angle=0), legend.title=element_text(size=18, angle=0),
        legend.text=element_text(size=16, angle = 0))

Fig_7

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_7.eps",Fig_7, width=12.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_7.pdf",Fig_7, width=12.16, height=7.45, limitsize = FALSE)

```

```{r Case of Schools - Fig: 8, 9, 10, 12, 13 and 14, echo=FALSE, fig.align="center", fig.height=8, fig.width=12, message=FALSE, warning=FALSE}
library(rpart)
library(dplyr)
library(readr)
library(rattle)
library(ggplot2)
library(reshape2)
library(ggforce)
library(gridExtra)
library(ggthemes)
setwd("~/Documents/GitHub/HE_Chile")

# Load --------------------------------------------------------------------

Base_Colegios=
  read_rds("Base_Colegios.RDS") %>% 
  mutate(
    COD_DEPE2=
      case_when(COD_DEPE2 == "Municipal" ~ "Public",
                COD_DEPE2 == "P.Sub" ~ "Semi-Private",
                COD_DEPE2 == "P.Pag" ~ "Private"),
     COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")))

Tipo_Institucion=read_rds("Tipo_Institucion.RDS") %>% 
  mutate(AGNO_EGRESO=AGNO_EGRESO+1) %>% 
  mutate(
    TIPO_INST_2=
      case_when(
        TIPO_INST_2 == "Centros de Formación Técnica" ~ "TTC",
        TIPO_INST_2 == "Institutos Profesionales" ~ "PI",
        TIPO_INST_2 == "Universidades CRUCH" ~ "CRUCH",
        TIPO_INST_2 == "Universidades Privadas" ~ "Non-CRUCH"
      ),
    TIPO_INST_2 = factor(TIPO_INST_2, levels=c("CRUCH", "Non-CRUCH", "TTC","PI"))
  )

# fig 8 -------------------------------------------------------------------

Base_Colegios_Tree=
  Base_Colegios %>% filter(NO_MATRICULADO+MATRICULADO>19) %>% 
  mutate(AGNO_EGRESO=ifelse(AGNO_EGRESO %in% paste(2007:2011),"<2011",">2011")) %>% 
  rename(HGPA=NEM,Admission=AGNO_EGRESO,Dependence=COD_DEPE2)

tree=rpart(TASA~Admission + Dependence + HGPA,data=Base_Colegios_Tree)
fancyRpartPlot(tree,type=3,palettes="Greys",sub=NULL)

# fig 9 -------------------------------------------------------------------

Count.Depe=
  acast(Base_Colegios %>% mutate(AGNO_EGRESO=as.numeric(AGNO_EGRESO)+2007),
        AGNO_EGRESO ~ COD_DEPE2, length) %>% 
  as.data.frame() %>% 
  melt(variable.name="DEPE",value.name="COUNT") %>% 
  mutate(AGNO=rep(2008:2020,3)) %>%
  mutate(DEPE = factor(DEPE, levels=c("Public", "Semi-Private", "Private")))

Fig_9 <- ggplot(Count.Depe) + aes(x=AGNO,y=COUNT,group=DEPE) +
  geom_line(linetype = "3313")+
  geom_point(aes(shape=factor(DEPE)), size=4)+ 
  geom_text(aes(label=COUNT),vjust=-0.9,hjust=0,size=4, color="black",check_overlap=TRUE,nudge_y=0.5) +
  scale_color_discrete("Dependence") +
  theme_bw() + 
  scale_x_continuous(breaks=2008:2020) +
  scale_y_continuous(limits=c(0,1800)) + 
  labs(x="Admission Process",y="Number of High Schools", shape="Dependence") +
  theme(legend.position="top",
        axis.title.y.right = element_blank(),
        plot.title = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),
        axis.title.x = element_text(size = 18, vjust=-0.5),
        axis.text.y = element_text(size=15, angle = 0),
        axis.title.y = element_text(size=18, angle=90),
        axis.text.x = element_text(size=14, angle=90)) +
  scale_shape_manual(values = c('Public'=15, 'Semi-Private'=17, 'Private'=16))

Fig_9

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_9.eps",Fig_9, width=10.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_9.pdf",Fig_9, width=10.16, height=7.45, limitsize = FALSE)

# fig 10 ------------------------------------------------------------------

Fig_10 <- ggplot(Tipo_Institucion %>% filter(TIPO_INST_2!="No Matriculado")) + 
  aes(x=AGNO_EGRESO,y=N_distintas,group=TIPO_INST_2) +
  geom_line(linetype = "3313")+
  geom_point(aes(shape=factor(TIPO_INST_2)), size=4)+ 
  geom_text_repel(aes(label=N_distintas)) +
  scale_color_discrete("Institutions") + #labels=c("All")
  theme_bw() + 
  scale_x_continuous(breaks=2008:2020) +
  scale_y_continuous(limits=c(0,80)) +
  labs(x="Admission Process", y="Number of Higher Education Institutions", shape="Institution") +
  guides(color=guide_legend(ncol=4)) +
  theme(legend.position="top",
        axis.title.y.right = element_blank(),
        plot.title = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),
        axis.title.x = element_text(size = 18, vjust=-0.5),
        axis.text.y = element_text(size=15, angle = 0),
        axis.title.y = element_text(size=18, angle=90),
        axis.text.x = element_text(size=14, angle=90)) +
  scale_shape_manual(values = c('CRUCH'=15, 'Non-CRUCH'=17, 'TTC'=16, 'PI'=8))

Fig_10

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_10.eps",Fig_10, width=10.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_10.pdf",Fig_10, width=10.16, height=7.45, limitsize = FALSE)

# fig 15 ------------------------------------------------------------------

Base_Colegios_Hist=Base_Colegios %>% mutate(AGNO_EGRESO=AGNO_EGRESO %>% as.numeric() + 2007)

ggplot(Base_Colegios_Hist %>% filter(COD_DEPE2=="Public"), aes(TASA)) +
  facet_wrap(~AGNO_EGRESO,ncol=4) +
  labs(x="Enrolment rate",y="Density") +
  ggtitle("Public High School") + 
  geom_histogram(color="black",fill="grey",alpha=0.3) + 
  theme_clean() +
  theme(legend.position="top",
        axis.title.y.right = element_blank(),
        plot.title = element_text(size = 18, face = "bold",hjust = 0.5),
        legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),
        axis.title.x = element_text(size = 18, vjust=-0.5),
        axis.text.y = element_text(size=15, angle = 0),
        axis.title.y = element_text(size=18, angle=90),
        axis.text.x = element_text(size=14, angle=90))

# fig 16 ------------------------------------------------------------------

ggplot(Base_Colegios_Hist %>% filter(COD_DEPE2=="Semi-Private"), aes(TASA)) +
  facet_wrap(~AGNO_EGRESO,ncol=4) +
  labs(x="Enrolment rate",y="Density") +
  ggtitle("Semi-Private High School") + 
  geom_histogram(color="black",fill="grey",alpha=0.3) + 
  theme_clean() +
  theme(legend.position="top",
        axis.title.y.right = element_blank(),
        plot.title = element_text(size = 18, face = "bold",hjust = 0.5),
        legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),
        axis.title.x = element_text(size = 18, vjust=-0.5),
        axis.text.y = element_text(size=15, angle = 0),
        axis.title.y = element_text(size=18, angle=90),
        axis.text.x = element_text(size=14, angle=90))

# fig 17 ------------------------------------------------------------------

ggplot(Base_Colegios_Hist %>% filter(COD_DEPE2=="Private"), aes(TASA)) +
  facet_wrap(~AGNO_EGRESO,ncol=4) +
  labs(x="Enrolment rate",y="Density") +
  ggtitle("Private High School") + 
  geom_histogram(color="black",fill="grey",alpha=0.3) + 
  theme_clean() +
  theme(legend.position="top",
        axis.title.y.right = element_blank(),
        plot.title = element_text(size = 18, face = "bold",hjust = 0.5),
        legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),
        axis.title.x = element_text(size = 18, vjust=-0.5),
        axis.text.y = element_text(size=15, angle = 0),
        axis.title.y = element_text(size=18, angle=90),
        axis.text.x = element_text(size=14, angle=90))

```

```{r Individual Multinomial Logit, echo=FALSE, fig.align="center", fig.height=8, fig.width=15, message=FALSE, warning=FALSE}
reg <- read_rds("BASE_PAPER.RDS")
reg <- reg %>% rename(NOMBRE_REGION=nombre_region, NOMBRE_COMUNA=nombre_comuna)

#Multinomial Logit
reg = reg %>% mutate(AGNO_EGRESO = AGNO_EGRESO+1)
reg$AGNO_EGRESO = factor(reg$AGNO_EGRESO)
reg <- reg %>% filter(!is.na(NEM)) %>% filter(!is.na(NOMBRE_COMUNA)) %>% filter(GEN_ALU!="0")
reg <- reg %>% filter(JOVEN==1)

reg <- within(reg, AGNO_EGRESO <- relevel(AGNO_EGRESO, ref = "2008"))
reg <- within(reg, EDAD <- relevel(EDAD, ref = "17"))
reg <- within(reg, GEN_ALU <- relevel(GEN_ALU, ref = "2"))
reg <- reg %>% mutate(NEM=NEM/100)

reg = reg %>% mutate(TIPO_INST_2=ifelse(is.na(TIPO_INST_2),"No-Matriculado", TIPO_INST_2))

reg <- reg %>% mutate(
    ESTADO =
      case_when(
        TIPO_INST_2 == "Centros de Formación Técnica" ~ "CFT e IP",
        TIPO_INST_2 == "Institutos Profesionales" ~ "CFT e IP",
        TIPO_INST_2 == "Universidades CRUCH" ~ "U. CRUCH",
        TIPO_INST_2 == "Universidades Privadas" ~ "U Privada No-CRUCH",
        TIPO_INST_2 == "No-Matriculado" ~ "No Matriculado"
      )
  )

reg$ESTADO = factor(reg$ESTADO)
reg$ESTADO <- relevel(reg$ESTADO, ref = "No Matriculado")

require(foreign)
require(nnet)
require(ggplot2)
require(reshape2)

reg <- within(reg, ESTADO <- relevel(ESTADO, ref = "No Matriculado"))
reg = reg %>% filter(EDAD %in% c('16','17','18','19'))
reg <- within(reg, COD_DEPE2 <- relevel(COD_DEPE2, ref = "Municipal"))

m_logit <- multinom(ESTADO ~ EDAD  + NEM + GEN_ALU + 
                 COD_DEPE2 + factor(PostBeca) + factor(PERCENTIL) + factor(NOMBRE_REGION) + AGNO_EGRESO, data = reg)

coeficientes = coef(m_logit)

#Coeficientes of the table.
#command Summary takes too long

View(coeficientes)

#Performance of the model:
library(performance)
r2_mcfadden(m_logit)

a = summary(m_logit)
z <- a$coefficients/a$standard.errors
z

p <- (1 - pnorm(abs(z), 0, 1)) * 2
p

library(stats)
logLik(m_logit)
modelChi <- m_logit$null.deviance - m_logit$deviance
pseudo.R2 <- modelChi / m4$null.deviance
pseudo.R2

lmtest::lrtest(m_logit)

```

```{r Extra: Mosaic Region-Dependence, echo=FALSE, fig.height=7, fig.width=15, message=FALSE, warning=FALSE}
Data <- read_rds("BASE_PAPER.RDS")
Data <- Data %>% filter(JOVEN=="1")
Data <- Data %>% rename(COD_REG_ALU=codigo_region,Region=nombre_region)

df <- Data %>% filter(!is.na(COD_REG_ALU)) %>% filter(COD_REG_ALU>0)

df <- df %>%
  group_by(Region,COD_REG_ALU,COD_DEPE2) %>%
  summarise(count = n()) %>%
  mutate(Region.count = sum(count),
         prop = count/sum(count)) %>%
  ungroup()

#df$TIPO_INST_3[is.na(df$TIPO_INST_3)]="No Matriculado"
#df <- df %>% arrange(COD_REG_ALU)
df$COD_REG_ALU[df$COD_REG_ALU==15]=0.5
df$COD_REG_ALU[df$COD_REG_ALU==16]=7.5
df$COD_REG_ALU[df$COD_REG_ALU==14]=9.5
df$COD_REG_ALU[df$COD_REG_ALU==13]=5.5

df <- df %>% mutate(Region = ifelse(Region=="METROPOLITANA DE SANTIAGO", "RM",
                                  ifelse(Region=="VALPARAISO","VALPO",
                                         ifelse(Region =="BIOBIO", "BBIO",
                                                ifelse(Region=="LA ARAUCANIA", "ARAUC",
                                                       ifelse(Region=="LIBERTADOR GENERAL BERNARDO O'HIGGINS", "LGBO",
                                                              ifelse(Region=="LOS LAGOS","LAGOS",
                                                                     ifelse(Region=="COQUIMBO", "COQ",
                                                                            ifelse(Region=="ANTOFAGASTA","ANTOF",
                                                                                   ifelse(Region=="LOS RIOS", "RIOS",
                                                                                          ifelse(Region=="TARAPACA", "TPCA",
                                                                                                 ifelse(Region=="ATACAMA","ATCMA",
                                                                                                        ifelse(Region =="ARICA Y PARINACOTA", "AYP",
                                                                                                               ifelse(Region=="MAGALLANES Y DE LA ANTARTICA CHILENA","MAG", ifelse(Region=="AYSEN DEL GENERAL CARLOS IBANEZ DEL CAMPO", "AYSEN", Region)))))))))))))))

df <- df %>% mutate(COD_DEPE2 = ifelse(COD_DEPE2=="Municipal","Public",
                                                     ifelse(COD_DEPE2=="P.Pag", "Private",
                                                            ifelse(COD_DEPE2=="P.Sub","Semi-Private",COD_DEPE2))))

Fig_11=ggplot(df,
       aes(x = reorder(Region,COD_REG_ALU), y = prop, width = Region.count, fill =COD_DEPE2 )) + labs(fill="Dependence") +
  geom_bar(stat = "identity", position = "fill", colour = "black") +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), position = position_stack(vjust = 0.5), size=3, angle=90) + 
  facet_grid(~reorder(Region,COD_REG_ALU), scales = "free_x", space = "free_x") +
  scale_fill_brewer(palette = "Pastel1") + 
  theme_bw() + theme(strip.text.x = element_text(size = 10, colour = "black", angle = 90), 
                     legend.position = "top", axis.text = element_blank(), 
                     legend.text = element_text(size=14), legend.title = element_text(size=16),
                     axis.ticks = element_blank()) + 
  labs(y="", x="")

Fig_11

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_11.eps",Fig_11, width=12.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_11.pdf",Fig_11, width=12.16, height=7.45, limitsize = FALSE)

```

```{r Extra: Rate by Region, echo=FALSE, fig.height=7, fig.width=15, message=FALSE, warning=FALSE}

Data <- read_rds("BASE_PAPER.RDS")
Data <- Data %>% filter(JOVEN=="1")
Data <- Data %>% rename(COD_REG_ALU=codigo_region,Region=nombre_region)

#admission process = agno_egreso + 1

matriculado_region <- Data %>% 
  filter(JOVEN=="1") %>%
  group_by(COD_REG_ALU, Region, MATRICULADO, TIPO_INST_1) %>% summarise(N=n())

matriculado_region=dcast(matriculado_region,COD_REG_ALU + Region ~ TIPO_INST_1 + MATRICULADO)
matriculado_region <- matriculado_region %>% rename(Universidades=Universidades_1 , CFT=`Centros de Formación Técnica_1` ,IP=`Institutos Profesionales_1`)
matriculado_region <- matriculado_region %>% rename(No_Matriculados=NA_0)
matriculado_region <- matriculado_region %>% mutate(Total_Egresados=Universidades+CFT+IP+No_Matriculados)
matriculado_region <- matriculado_region %>% mutate(Total_Matriculados=Universidades+CFT+IP)
matriculado_region <- matriculado_region %>% filter(!is.na(COD_REG_ALU)) %>% filter(COD_REG_ALU>0)
matriculado_region <- matriculado_region %>% mutate(Tasa_Ingreso = Total_Matriculados/Total_Egresados*100)

matriculado_region <- matriculado_region %>% mutate(Region = ifelse(Region=="METROPOLITANA DE SANTIAGO", "RM",
                                  ifelse(Region=="VALPARAISO","VALPO",
                                         ifelse(Region =="BIOBIO", "BBIO",
                                                ifelse(Region=="LA ARAUCANIA", "ARAUC",
                                                       ifelse(Region=="LIBERTADOR GENERAL BERNARDO O'HIGGINS", "LGBO",
                                                              ifelse(Region=="LOS LAGOS","LAGOS",
                                                                     ifelse(Region=="COQUIMBO", "COQ",
                                                                            ifelse(Region=="ANTOFAGASTA","ANTOF",
                                                                                   ifelse(Region=="LOS RIOS", "RIOS",
                                                                                          ifelse(Region=="TARAPACA", "TPCA",
                                                                                                 ifelse(Region=="ATACAMA","ATCMA",
                                                                                                        ifelse(Region =="ARICA Y PARINACOTA", "AYP",
                                                                                                               ifelse(Region=="MAGALLANES Y DE LA ANTARTICA CHILENA","MAG", ifelse(Region=="AYSEN DEL GENERAL CARLOS IBANEZ DEL CAMPO", "AYSEN", Region)))))))))))))))

matriculado_region <- matriculado_region %>% rename(codigo_region=COD_REG_ALU)

x = generar_regiones(mapa = chilemapas::mapa_comunas)
x$codigo_region = as.numeric(x$codigo_region)
x <- left_join(x, matriculado_region)
x <- x %>% mutate(Zoom=ifelse(codigo_region<8 | codigo_region==15 | codigo_region==16 | codigo_region==13,1,0))

x <- x %>% mutate(Porc = paste0(round(Tasa_Ingreso,1),"%"))

bar = ggplot(data=x, aes(x=reorder(factor(Region),Tasa_Ingreso),y=Tasa_Ingreso, fill=Tasa_Ingreso))  + coord_flip() + theme_minimal() + geom_bar(stat="identity") + labs(x="Region",y="Enrolment Rate") + theme(legend.position = "none", axis.title = element_text(size=17), axis.text.y = element_text(size=14), axis.text.x = element_text(size=14)) + geom_text(aes(label = Porc), position = position_stack(vjust = 0.5), size = 5, color="black") + scale_y_continuous(expand = expansion(mult = c(0, .1)))+scale_fill_distiller(palette = "RdBu", direction=1,limits=c(40,63))
bar

```